---
title: "Lab 10"
author: " "
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

For this lab we will carry out the paper helicopter experiment that you helped design for Lab 9 and HW 4. We will use a randomized complete block design, with subsamples.

Using the following guide  ([https://www.wikihow.com/Create-a-Paper-Helicopter](https://www.wikihow.com/Create-a-Paper-Helicopter))
each student (and block) will make a total of 4 helicopters.

- `wing_length`: 8 cm and 12 cm
- `wing_width`: 3 cm and 6 cm

Run the following code and update `set.seed()` with your birthday `set.seed(MMDD)`to determine the order that you make and drop the helicopters. 

```{r}
set.seed(1015)
tibble(EU = 1:4,
       trt = sample(paste('length ', rep(c(8, 12),2), '; width ', rep(c(3, 6), each = 2), sep = '')))
```

There is no official protocol for body length, but be consistent across your 4 helicopters.

To understand the observational error (differences in drop time for the same helicopter), each helicopter will be dropped and timed three times. Helicopters should be dropped from the bridge off of Wilson Hall. Drop them with an arm extended over the bridge over the where the sidewalk crosses under the bridge.

Please submit the following a CSV file with the following structure, where time in seconds contains the time before the helicopter touches the ground (replace the 4s in the dataset below).

```{r}
tibble(EU = rep(1:4, each = 3),
       trt = rep(paste('length ', rep(c(8, 12),2), '; width ', rep(c(3, 6), each = 2), sep = ''), each = 3),
       OU = rep(1:3, 4),
       time = 4)
```

Hint, you can use `write_csv()` in R to pass an R data frame to a CSV file.
