---
title: "Lab 10"
author: " "
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

For this lab we will carry out the paper helicopter experiment that you helped design for Lab 9 and HW 4. We will use a randomized complete block design, with subsamples.

Using the following guide  ([https://www.wikihow.com/Create-a-Paper-Helicopter](https://www.wikihow.com/Create-a-Paper-Helicopter))
each student (and block) will make a total of 4 helicopters.

- `wing_length`: 8 cm and 12 cm
- `wing_width`: 3 cm and 6 cm

Run the following code and update `set.seed()` with your birthday `set.seed(MMDD)`to determine the order that you make and drop the helicopters. 

```{r}
set.seed(1015)
tibble(EU = 1:4,
       trt = sample(paste('length ', rep(c(8, 12),2), '; width ', rep(c(3, 6), each = 2), sep = '')))
```

There is no official protocol for body length, but be consistent across your 4 helicopters.

To understand the observational error (differences in drop time for the same helicopter), each helicopter will be dropped and timed three times.

Please submit the following a CSV file with the following structure, where time contains the time before the helicopter touches the ground.

```{r}
tibble(EU = rep(1:4, each = 3),
       trt = rep(paste('length ', rep(c(8, 12),2), '; width ', rep(c(3, 6), each = 2), sep = ''), each = 3),
       OU = rep(1:3, 4),
       time = 4)
```

Hint, you can use `write_csv()` in R to pass an R data frame to a CSV file.
