---
title: "Lab5"
author: ""
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(see)
library(arm)
```

For this lab we will be starting to think about analyzing our airplane data. Clean the `Airplane` dataset and recreate a figure similar to lab 2.


```{r, message = F, warning = F}
airplane <- read_csv("https://raw.githubusercontent.com/stat441/Labs/main/airplane_clean.csv")

airplane_wide <- airplane %>% 
  mutate(value = feet_dec) %>% 
  dplyr::select(-feet_dec) %>% 
  pivot_wider(names_from = name, values_from = value)

```


## Data Visualization

```{r, fig.align = 'center', fig.cap = 'Distance traveled for paper airplane in STAT 441 / 541 experiment. Numbers represent unique paper airplane maker/throwers.'}
airplane %>% 
  ggplot(aes(y = feet_dec, x = name, label = id)) +  
  geom_violinhalf() + 
  geom_boxplot(width=0.1) + 
  geom_text(position = position_jitter(seed = 1)) + 
  theme_minimal() + 
  ylab('Distance traveled (feet)') + 
  xlab('Airplane Type') + 
  ggtitle('Airplane Distance from STAT441/541 Experiment')
```


#### 1. 

Write out the statistical model suggested implied by the following code. 

```{r}
lm(feet_dec ~ name, data = airplane) %>% display()
```


#### 2. 

Write out the statistical model suggested implied by the following code.

```{r}
lm(feet_dec ~ name - 1, data = airplane) %>% display()
```


#### 3.

Do the statistical models in Q1 and Q2 account for the blocking structure of our designed experiment? If not, evaluate a model to include this factor using the reference case specification of Q1. Note: you don't need to write out the model notation, just fit this model.

```{r}
display(lm(feet_dec ~ name + factor(id), data = airplane))
```

#### 4.

Analyze the data using a paired t-test

```{r}
#t.test(x = , y = , paired = T) 
```


#### 5.

Analyze the data using a t-test on the difference (Dart - Glider) for each participant


#### 6.

Which of the analyses Q1 - Q5 provide the same inferences from the experiment?

