---
title: "Lab 7"
author: " "
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```


Every experimental design starts with a pre-randomized design plan that will:

- List out all possible treatment levels (could come from one or more treatment factors)
- Decide number of replicates per treatment level

For a completely randomized design (CRD) plan, assignment of treatments to EUs is done completely at random and preserves the desired replication.

---

#### Question 1: experimental plan

Consider our pizza dough example from class with three times allowing the dough to rise. Write out the pre-design plan (treatment levels & number of replicates). You can decide the number of replicates.

---

A randomization procedure is valid if:

 - all permutations of treatment labels are equally likely
 - it preserves the desired replication
 
This is not the same as saying every EU has the same chance to be assigned a given treatment

--

#### Question 2: valid randomization?

Would the following procedure be a valid randomization, why or why not?

```{r}
set.seed(03082022)
trt_levels <- 3
replicates <- 5
EUs <- trt_levels * replicates

data.frame(EUs = 1:EUs, trt = sample(3,15, replace = T)) %>% kable()
```

---

Consider a design that with two treatment levels and 3 replicates for each replicate. Assume randomization is determined by a coin flip for each participant - until the quota for given treatment level has been met.

Consider the following randomization and associated probabilities

```{r, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| EUs        | 1 | 2  | 3| 4 | 5  | 6 |
|------|:--------:|------:|------:|------|:--------:|------:|
| TRT    | 1      | 2 | 2 |1| 1| 2|
| Prob | 1/2  |1/2|1/2 | 1/2| 1/2| 1 |
"
cat(tabl) 
```

for this randomization procedure, the last treatment procedure is completely determined. So the probability of this design would be $\left( \frac{1}{2} \right)^5$

---

#### Question 3: permutation probability

Calculate the probability of each permutation below. Does this result in a valid randomization procedure?

```{r, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
| EUs        | 1 | 2  | 3| 4 | 5  | 6 |
|------|:--------:|------:|------:|------|:--------:|------:|
| perm 1    | 1      | 2 | 2 |2| 1| 1|
| perm 2    | 2      | 2 | 2 |1| 1| 1|
| perm 3    | 1      | 2 | 2 |1| 1| 2|
"
cat(tabl) 
```

---

There are many valid ways to generate a CRD. The easiest is to use computer randomization:

1. Create a list of EUs from 1 to N
2. Generate random number for each EU
3. Sort by EUs by random number
4. Assign first $r_1$ to treatment 1, assign next $r_2$ to treatment 2...

Note steps 1 - 3 can be done with sample in `R`.

---

#### Question 4: design randomization

Write code to create a valid randomization for your experimental settings in Question 1.

---

Treatments do not have to be equally replicated for a design to be called a CRD, its a CRD due to the randomization procedure.

CRDS are recommended when EUs are fairly homogeneous

Nuisance factors are either:

- Held constant
- Measured and adjusted for with analysis of covariance

