---
title: "Lab8"
author: ""
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

For this lab we will explore ANCOVA models using simulated data. Consider the motivation presented in class where the experimental units are piglets of different weights.

### Q1
Consider the 60 "fake" piglets and write code to add a column titled `treatment` to the `fake_pigs` tibble. Use a CRD to allocate 20 EUs to treatments 1, 2, and 3.

```{r}
set.seed(03212022)
total_pigs <- 60
fake_pigs <- tibble(piglet = 1:total_pigs, piglet_weight = runif(total_pigs, min = 10, max = 30))
```

### Q2

Consider the following statistical model:
$$Y_{ij} = \tau_i + x_{ij} \beta + E_{ij}$$
where $Y_{ij}$ is the weight after a study for the $j^{th}$ piglet in the $i^{th}$ treatment, $\tau_i$ is the treatment effect associated with treatment $i$ (when holding $\mu = 0$ for identifiability), $x_{ij}$ is the starting weight for the $ij_{th}$ piglet, $\beta$ is effect associated with the starting weight, and $E_{ij}$ is the error term in the model, where $E_{ij} \sim N(0, \sigma^2).$

Data has been simulated for each of the treatments using values below for $\tau_1$, $\tau_2$, $\tau_3$, $\beta$, and $\sigma$ simulate values for $Y$. Create a figure that shows Y and includes indicators (color / shapes) for the different treatments. (Note: you'll need to make sure you add a column titled `treatment` in your `fake_pigs` tibble and remove the eval = F tag on the R code chunk below.)

```{r}
tau1 <- 10
tau2 <- 15
tau3 <- 20
beta <- 1
sigma <- 2
```
```{r, eval = F}
X_matrix <- model.matrix(~factor(treatment) - 1 + piglet_weight, data = fake_pigs)
param_vec <- c(tau1, tau2, tau3, beta)
Y <- rnorm(total_pigs, mean = X_matrix %*% param_vec, sd = sigma)
fake_pigs <- fake_pigs %>% mutate(Y = Y)
```

### Q3

Use `lm` to fit the model spelled in in Q2 to the sythetic data. Print the model output here. Use the cell means specification.

### Q4

Now consider the model in Q2, does the treatment effect differ depending on the starting weight? Why or why not?

### Q5

Consider the figure below and explain how the initial piglet weights and the treatment factors influence the final weights.

```{r, echo = F}
fake_pigs <- fake_pigs %>% 
  mutate(treatment = factor(rep(1:3, total_pigs / 3))) # not a valid CRD
X_matrix <- model.matrix(~factor(treatment) * piglet_weight -1, data = fake_pigs)
param_vec <- c(tau1, tau2, tau3, beta, -1, -.6)
new_Y <- rnorm(total_pigs, mean = X_matrix %*% param_vec, sd = sigma)
fake_pigs <- fake_pigs %>% mutate(Y2 = new_Y)

fake_pigs %>% 
  ggplot(aes(y = Y2, x = piglet_weight, color = treatment)) +
  geom_point() + 
  geom_smooth(method = 'lm', formula = 'y~x') + 
  theme_bw() + xlab('Initial Piglet Weight') +
  ylab('Final piglet weight') +
  ggtitle('Piglet weights') +
  ylim(0, max(new_Y)) + xlim(0, max(fake_pigs$piglet_weight))
```


### Q6 (541 only)

Fit a model to the data set created for Q5 that includes an interaction term. Interpret the parameters in this model - for this you don't need to talk about the actual values, just the meaning of the coefficients.
